\dppTitle

### 2014-06-01 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Working on dpp ms revisions.

\change{
\msb has been used to address biogeographical questions in variety of empirical
systems. Some examples include
(1) whether the rise of the Isthmus of Panama caused co-divergence among
species of echinoids co-distributed across the Pacific and Atlantic sides of
the isthmus \cite{Hickerson2006}, 
(2) if an historical seaway across the Baja Peninsula caused co-divergence
across species of squamates and mammals co-distributed both north and south of
the putative seaway \cite{Leache2007},
(3) if species of gall-wasps and their associated parasitoids share divergences
across putative glacial refugia \cite{Stone2012}, and
(4) whether repeated coalescence and fragmentation of the oceanic Islands of
the Philippines during Pleistocene sea-level fluctuations caused
diversification of vertebrate taxa distributed across the islands
\cite{Oaks2012}.
Such applications of the method often result in strong posterior support for
co-divergence among some of the taxa investigated (e.g.,
\cite{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012}).
}

For priors on divergence-time and demographic parameters, \msb uses continuous
uniform probability distributions.
\change{
This causes divergence models with more divergence-time parameters to integrate
over a \emph{much} greater parameter space with low likelihood yet high prior
density, which can result in small marginal likelihoods relative to models with
fewer divergence-time parameters \cite{Jeffreys1939,Lindley1957}.
Given that the marginal likelihood of a model weighted by its prior is what
determines its posterior probability, this can cause support for models with
fewer divergence events \cite{Oaks2012,Oaks2014reply}.
% The use of marginal likelihoods to estimat model probabilities provides a
% ``natural'' penalty for overparameterization, and can be a great strength of
% the Bayesian approach.
This is not a critique of Bayesian model choice in general; comparing models by
their marginal likelihoods provides a ``natural'' penalty for
overparameterization and can be a great strength of the Bayesian approach.
However, given the sensitivity of marginal likelihoods to the prior, care is
needed when selecting prior distributions \cite{Lindley1957}.
Selecting distributions that will often place high prior density in large
regions of parameter space with low likelihood can lead to small marginal
likelihoods of parameter-rich models even if they are correct.
}

Furthermore, \msb uses a discrete uniform prior over the number of divergence
events $1,2,\ldots,\npairs{}$.
Because there are many more possible assignments of populations pairs to
intermediate numbers of divergence events, this imposes a prior on divergence
models that puts most of the prior mass on models with either very few or very
many divergence-time parameters (see Figure 5 of \cite{Oaks2012}; for brevity
I will refer to this prior as ``U-shaped'').
Given that models with many divergence events can have small marginal
likelihoods due to the uniform priors on divergence-time parameters, the
U-shaped prior will effectively create a strong prior preference for models
with very few divergence events.

\change{
Recently, Oaks et al.\ \cite{Oaks2012,Oaks2014reply} found via simulation that
\msb will often strongly support models with a small number of divergence
events shared among taxa, even when divergences were random over broad
timescales.}
They suggested this behavior was due to the combination of uniform priors on
parameters causing small marginal likelihoods of richer models and the
U-shaped prior on divergence models.
Hickerson et al.\ \cite{Hickerson2013} suggested the problem is caused by
sampling error, and proposed as a solution an approximate-Bayesian model
averaging approach that samples over empirically informed uniform priors.
However, Oaks et al.\ \cite{Oaks2014reply} evaluated the approach proposed by
Hickerson et al.\ \cite{Hickerson2013} using simulations and found that it did
not mitigate the method's propensity to incorrectly infer clustered
divergences, and often preferred priors that excluded the true values of the
model's parameters.
\change{
Here, I describe a new method that successfully mitigates spurious inference of
co-divergence while avoiding negative side effects of empirically informed
uniform priors.}

@@

\begin{linenomath}
The prior term for the demographic parameters, $p(\demographicParamVector)$,
expands out into its components and as a product over the \npairs{}
pairs of populations
\begin{equation}
    p(\demographicParamVector) =
    \prod_{i=1}^{\npairs{}}
    p(\ancestralTheta{i})
    p(\descendantTheta{1}{i})
    p(\descendantTheta{2}{i})
    p(\bottleScalar{1}{i})
    p(\bottleScalar{2}{i})
    p(\bottleTime{i})
    p(\migrationRate{i}).
    \label{eq:demographicPrior}
\end{equation}
\change{
The priors for the demographic parameters are
$\ancestralTheta{} \sim Gamma(\gammaShape{\ancestralTheta{}},
\gammaScale{\ancestralTheta{}})$,
$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
\gammaScale{\descendantTheta{}{}})$,
$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
\gammaScale{\descendantTheta{}{}})$,
$\bottleScalar{1}{} \sim Beta(\betaA{\bottleScalar{}{}},
\betaB{\bottleScalar{}{}})$,
$\bottleScalar{2}{} \sim Beta(\betaA{\bottleScalar{}{}},
\betaB{\bottleScalar{}{}})$,
$\bottleTime{} \sim U(0, 1)$,
and
$\migrationRate{} \sim Gamma(\gammaShape{\migrationRate{}},
\gammaScale{\migrationRate{}})$,
where the hyper-parameters of each prior distribution can be specified by
the investigator.
By default, \ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}
share the same prior (i.e., $\gammaShape{\ancestralTheta{}} =
\gammaShape{\descendantTheta{}{}}$ and $\gammaScale{\ancestralTheta{}} =
\gammaScale{\descendantTheta{}{}}$), but a separate gamma-distributed prior can
be assigned to \ancestralTheta{}.
Also, the \bottleScalar{1}{}, \bottleScalar{2}{}, and \migrationRate{}
parameters are optional (i.e., the investigator can assume that there has been
no migration between populations of each pair and/or the population size of
each descendant population has been constant through time).}
\end{linenomath}

@@

\subsubsection{Ordering of taxon-specific summary statistics}
As alluded to in the model description, \msb does not maintain the order of the
taxon-specific summary statistics \alignmentSS{}{} within each \ssVector{}.
Rather, the summary statistics are re-ordered by descending values of average
pairwise differences between the descendant populations
($\pi_b$) \cite{NeiLi1979,Huang2011}.
\change{
As a result, the summary statistics being compared when calculating the
Euclidean distance between the observed data and each simulated dataset often
represent sequence alignments of \emph{different} taxon pairs and/or loci.
More specifically, the summary statistics calculated from the observed sequence
alignments are being compared to summary statistics calculated from datasets
simulated with potentially \emph{different} 
(1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
(2) length of alignments,
(3) numbers of loci (\nloci{}),
(4) \hky model parameters (\hkyModel{}{}),
(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
and
(6) ploidy multipliers (\ploidyScalar{}{}).}

This has the advantage of reducing the sample space of the number of
divergence models \divTimeIndexVector, but there are at least two disadvantages.
First, additional information in the data is lost.
By discarding the identity of the \npairs{} pairs of populations, all
pair-specific information about the amount of data (e.g., the number of gene
copies collected from each of the populations [\popSampleSize{1}{} and
\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
taxon- and locus-specific parameters (\hkyModel{}{},
\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
\locusMutationRateScalar{}) is lost.
Second, the results are more difficult to interpret, because they can no
longer be directly associated to the taxa under study.

In the original descriptions of the \msb model \cite{Hickerson2006,Huang2011},
this re-ordering is justified by the fact that the expected value of $\pi_b$ is
unrelated to sample size \popSampleSize{1}{} and \popSampleSize{2}{} and thus
exchangeable among pairs.
\change{
This is incorrect for two reasons.}
First, the entire vector of summary statistics \alignmentSS{}{} for each pair
of populations is re-ordered across pairs, which implies that the justification
for re-ordering $\pi_b$ applies to all the statistics within each
\alignmentSS{}{}.
\change{
However, the expectations for statistics that estimate gross diversity (e.g.,
$\pi$ and $\theta_W$) are not independent of sample size for structured
populations (e.g., the divergent pairs of populations modeled by \msb), and
other statistics are not independent of sample size in general (e.g.,
$SD(\pi-\theta_W)$).}
Second, and more importantly, having the same expectation does not ensure
random variables are exchangeable.
Rather, for variables to be exchangeable their marginal distributions must be
the same (i.e., they must be identically distributed).
None of the summary statistics used by \msb, including $\pi_b$, are identically
distributed across different sample sizes.
\change{
More generally, the simulated alignments and their summary statistics are not
identically distributed if there is any variation among taxa or loci in the
(1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
(2) length of alignments,
(3) numbers of loci (\nloci{}),
(4) \hky model parameters (\hkyModel{}{}),
(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
or
(6) ploidy multipliers (\ploidyScalar{}{}).}
\change{
Whenever such variation is present (i.e., nearly all empirical applications),
the taxon-specific summary statistics \alignmentSS{}{} are not exchangeable,
and the reshuffling of the summary statistic vectors in not mathematically
valid.

The magnitude of the affect of this violation of exchangeability is not known.
Huang et al.\ \cite{Huang2011} demonstrated that the reordering of the summary
statistic vectors can greatly increase the method's tendency to infer a single
divergence event.
By definition, if the summary statistic vectors were exchangeable, the
reordering would not change the likelihood or posterior (barring sampling
error).
Thus, the results of Huang et al.\ \cite{Huang2011} suggest the reordering of
the statistics can introduce bias.
% Thus, the theoretical basis for this reshuffling is questionable.
% It may be justifiable as an additional coarsening of the data,
% by removing the information associated with the identity of the
% population pairs.
% However, I do not provide a formal proof that this re-ordering of the pairs
% does not introduce bias.
% One can imagine situations in which the sampling intensity (i.e., the
% number of gene copies, loci, and locus length) is highly skewed across
% pairs and/or there is large heterogeneity in mutational processes
% (e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
% In such cases, it seems possible that discarding this information could
% cause bias.
% Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
% is to minimize Euclidean distances between simulated datasets in which the true
% model has a single divergence event \cite{Huang2011}, and (2) the reordering
% increases the model's tendency to infer a single divergence event
% \cite{Huang2011}, it seems possible that this approach could be biasing the
% method \cite{Oaks2012}.
}

@@

For each set of 50,000 simulated datasets, I used the posterior estimates
to assess the model-choice behavior of each model.
I did this by assigning the 50,000 estimates of the posterior probability
of one-divergence event to 20 bins of width 0.05, and plotted
the estimated posterior probability of each bin against the proportion of
replicates in that bin with a true value consistent with one divergence
event \cite{Huelsenbeck2004,Oaks2012}.
\change{
Ideally, the estimated posterior probability of the one-divergence model
should estimate the probability that the one-divergence model is correct.
For large numbers of simulation replicates, the proportion of the replicates in
each bin for which the one-divergence model is true will approximate the
probability that the one-divergence model is the correct model.
Thus, if the method has the desirable behavior such that the estimated
posterior probability of the one-divergence model is an unbiased estimate of
the probability that the one-divergence model is correct, the points should
fall near the identity line.
For example, let us say the method estimates a posterior probability of 0.90
for 1000 simulated datasets.
If the method is accurately estimating the probability that the one-divergence
model is correct given the data, then the one-divergence model should be the
true model in approximately 900 of the 1000 replicates.
Any trend away from the identity line indicates the method is biased in the
sense that it is not accurately estimating the probability that the
one-divergence model is the correct model.

@@

\subsection{Validation analyses: Ordered divergence models}
The results show that the method performs similarily when sampling over ordered
models of divergence
(Figures S\ref{figValidationAccuracyDPPOrderedDPPOrdered} \&
S\ref{figValidationModelChoiceDPPOrderedDPPOrdered}).
This suggests that the method is not adversely affected by the increase
in the number of possible discrete models (from 22 unordered to 4140 ordered
models) when there are eight pairs of populations.
\change{
This is encouraging, because, as discussed above, the sorting of the summary
statistics calculated from sequence alignments is not valid for empirical
datasets in which there is variation in the number of individuals sampled per
population, generation times across taxa, nucleotide substitution parameters,
or the number, length, mutation rate, or ploidy of the loci sampled.
Thus, this sorting should not be performed for any empirical application
of the method.}

@@

\section{Discussion}
My results demonstrate that using alternative priors on parameters and
divergence models improved the behavior of the \msb model.
In the new implementation, model-choice estimation is more accurate and shows
greater robustness to model violations
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The original model is very sensitive to violations and, when present,
strongly over-estimates the probability of one-divergence event shared
across all taxa
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
\change{
When more appropriate priors are used for divergence-time and demographic
parameters, and either a Dirichlet-process or uniform prior applied across
divergence models, the model is less sensitive to violations, and, when they do
cause bias, the method is conservative and tends to underestimate the
probability of models with temporally clustered divergences
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).}

@@

\change{
I also urge caution when using \dppmsbayes due to the lack of theoretical
validation of Bayesian model choice when the full data are replaced by summary
statistics that are insufficient for discriminating across models under
comparison \cite{Robert2011}, which is certainly the case here.
Robert et al.\ \cite{Robert2011} demonstrated that ABC estimates of model
posterior probabilities can be biased when such across-model insufficient
statistics are used \cite{Robert2011}.
As recommended by Oaks et al.\ \cite{Oaks2012}, any results from the method
should be accompanied by assessments of prior sensitivity and power.
Until full-likelihood methods are available to assess the accuracy of ABC
estimates of divergence-model posterior probabilities, I encourage
investigators to view this method as a means of exploring their data rather
than a rigorous means of evaluating hypotheses.}

@@

\section{Conclusions}
I introduced a new approximate-Bayesian model-choice method for estimating
shared divergence histories across taxa using DNA sequence data.
The new method, \dppmsbayes, takes a non-parametric approach to the problem by
using a Dirichlet-process prior on the temporal distribution of divergences
across taxa.
The new method shows improved robustness, accuracy, and power compared to the
existing method, \msb.
\change{
Compared to \msb, the new approach better estimates posterior uncertainty,
which greatly reduces the chances of incorrectly estimating biogeographical
scenarios of shared divergence events.
This is important, because models of shared divergence events are often of
particular interest to researchers who employ these methods.
}
This new tool will allow evolutionary biologists to better leverage comparative
genetic data to assess the affects of regional and global biogeographical
processes on biodiversity.

>>>>

\change{
\msb has been used to address biogeographical questions in a variety of
empirical systems. Some examples include
(1) whether the rise of the Isthmus of Panama caused co-divergence among
species of echinoids co-distributed across the Pacific and Atlantic sides of
the isthmus \cite{Hickerson2006}, 
(2) if an historical seaway across the Baja Peninsula caused co-divergence
across species of squamates and mammals co-distributed both north and south of
the putative seaway \cite{Leache2007},
(3) if species of gall-wasps and their associated parasitoids share divergences
across putative glacial refugia \cite{Stone2012}, and
(4) whether repeated fragmentation of the oceanic Islands of the Philippines
during Pleistocene sea-level fluctuations caused diversification of vertebrate
taxa distributed across the islands \cite{Oaks2012}.
Such applications of the method often result in strong posterior support for
co-divergence among all or subsets of the taxa investigated (e.g.,
\cite{Barber2010,Carnaval2009,Chan2011,Hickerson2006,Leache2007,Plouviez2009,Stone2012,Voje2009,Oaks2012}).
}

For priors on divergence-time and demographic parameters, \msb uses continuous
uniform probability distributions.
\change{
This causes divergence models with more divergence-time parameters to integrate
over a \emph{much} greater parameter space with low likelihood yet high prior
density, which can result in small marginal likelihoods relative to models with
fewer divergence-time parameters \cite{Jeffreys1939,Lindley1957}.
Given that the marginal likelihood of a model weighted by its prior is what
determines its posterior probability, this can cause support for models with
fewer divergence events \cite{Oaks2012,Oaks2014reply}.
% The use of marginal likelihoods to estimat model probabilities provides a
% ``natural'' penalty for overparameterization, and can be a great strength of
% the Bayesian approach.
This is not a critique of Bayesian model choice in general; comparing models by
their marginal likelihoods provides a ``natural'' penalty for
over-parameterization and can be a great strength of the Bayesian approach.
However, given the sensitivity of marginal likelihoods to the prior, care is
needed when selecting prior distributions \cite{Lindley1957}.
Selecting distributions that will often place high prior density in large
regions of parameter space with low likelihood can lead to small marginal
likelihoods of parameter-rich models even if they are correct.
}

Furthermore, \msb uses a discrete uniform prior over the number of divergence
events $1,2,\ldots,\npairs{}$.
Because there are many more possible assignments of population pairs to
intermediate numbers of divergence events, this imposes a prior on divergence
models that puts most of the prior mass on models with either very few or very
many divergence-time parameters (see Figure 5 of \cite{Oaks2012}; for brevity
I will refer to this prior as ``U-shaped'').
Given that models with many divergence events can have small marginal
likelihoods due to the uniform priors on divergence-time parameters, the
U-shaped prior will effectively create a strong prior preference for models
with very few divergence events.

\change{
Recently, Oaks et al.\ \cite{Oaks2012,Oaks2014reply} found via simulation that
\msb will often strongly support models with a small number of divergence
events shared among taxa, even when divergences were random over broad
timescales.}
They suggested this behavior was due to the combination of uniform priors on
parameters causing small marginal likelihoods of richer models and the
U-shaped prior on divergence models.
Hickerson et al.\ \cite{Hickerson2013} suggested the problem was caused by
sampling error, and proposed as a solution an approximate-Bayesian model
averaging approach that samples over empirically informed uniform priors.
However, Oaks et al.\ \cite{Oaks2014reply} evaluated the approach proposed by
Hickerson et al.\ \cite{Hickerson2013} using simulations and found that it did
not mitigate the method's propensity to incorrectly infer clustered
divergences, and often preferred priors that excluded the true values of the
model's parameters.
\change{
Here, I describe a new method that successfully mitigates spurious inference of
co-divergence while avoiding negative side effects of empirically informed
uniform priors.}

@@

\begin{linenomath}
The prior term for the demographic parameters, $p(\demographicParamVector)$,
expands out into its components and as a product over the \npairs{}
pairs of populations
\begin{equation}
    p(\demographicParamVector) =
    \prod_{i=1}^{\npairs{}}
    p(\ancestralTheta{i})
    p(\descendantTheta{1}{i})
    p(\descendantTheta{2}{i})
    p(\bottleScalar{1}{i})
    p(\bottleScalar{2}{i})
    p(\bottleTime{i})
    p(\migrationRate{i}).
    \label{eq:demographicPrior}
\end{equation}
The priors for the demographic parameters are
$\ancestralTheta{} \sim Gamma(\gammaShape{\ancestralTheta{}},
\gammaScale{\ancestralTheta{}})$,
$\descendantTheta{1}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
\gammaScale{\descendantTheta{}{}})$,
$\descendantTheta{2}{} \sim Gamma(\gammaShape{\descendantTheta{}{}},
\gammaScale{\descendantTheta{}{}})$,
$\bottleScalar{1}{} \sim Beta(\betaA{\bottleScalar{}{}},
\betaB{\bottleScalar{}{}})$,
$\bottleScalar{2}{} \sim Beta(\betaA{\bottleScalar{}{}},
\betaB{\bottleScalar{}{}})$,
$\bottleTime{} \sim U(0, 1)$,
and
$\migrationRate{} \sim Gamma(\gammaShape{\migrationRate{}},
\gammaScale{\migrationRate{}})$,
\change{
where the hyper-parameters of each prior distribution can be specified by
the investigator.
By default, \ancestralTheta{}, \descendantTheta{1}{}, and \descendantTheta{2}{}
share the same prior (i.e., $\gammaShape{\ancestralTheta{}} =
\gammaShape{\descendantTheta{}{}}$ and $\gammaScale{\ancestralTheta{}} =
\gammaScale{\descendantTheta{}{}}$), but a separate gamma-distributed prior can
be assigned to \ancestralTheta{}.
Also, the \bottleScalar{1}{}, \bottleScalar{2}{}, and \migrationRate{}
parameters are optional (i.e., the investigator can assume that there has been
no migration between populations of each pair and/or the population size of
each descendant population has been constant through time).}
\end{linenomath}

@@

\subsubsection{Ordering of taxon-specific summary statistics}
As alluded to in the model description, \msb does not maintain the order of the
taxon-specific summary statistics \alignmentSS{}{} within each \ssVector{}.
Rather, the summary statistics are re-ordered by descending values of average
pairwise differences between the descendant populations
($\pi_b$) \cite{NeiLi1979,Huang2011}.
This has the advantage of reducing the sample space of possible divergence
models \divTimeIndexVector, but there are at least two disadvantages.
First, additional information in the data is lost.
By discarding the identity of the \npairs{} pairs of populations, all
pair-specific information about the amount of data (e.g., the number of gene
copies collected from each of the populations [\popSampleSize{1}{} and
\popSampleSize{2}{}], the number of loci, and the length of the loci), and the
taxon- and locus-specific parameters (\hkyModel{}{},
\mutationRateScalarConstant{}{}, \ploidyScalar{}{}, and
\locusMutationRateScalar{}) is lost.
\change{
Second, the results are more difficult to interpret, because divergence models
and parameter estimates cannot be directly associated to the taxa under study.
}

\change{
The re-ordering of the summary statistic vectors also has an important
implication for the ABC algorithm.
When calculating the Euclidean distance between the observed data and each
simulated dataset, the summary statistics being compared often represent
sequence alignments of \emph{different} taxon pairs and/or loci.
More specifically, the summary statistics calculated from the observed sequence
alignments are being compared to summary statistics calculated from datasets
simulated with potentially \emph{different} 
(1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
(2) length of alignments,
(3) numbers of loci (\nloci{}),
(4) \hky model parameters (\hkyModel{}{}),
(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
and
(6) ploidy multipliers (\ploidyScalar{}{}).}

In the original descriptions of the \msb method \cite{Hickerson2006,Huang2011},
this re-ordering is justified by the fact that the expected value of $\pi_b$ is
unrelated to sample size \popSampleSize{1}{} and \popSampleSize{2}{} and thus
exchangeable among pairs.
\change{
This is incorrect for two reasons.}
First, the entire vector of summary statistics \alignmentSS{}{} for each pair
of populations is re-ordered across pairs, which implies that the justification
for re-ordering $\pi_b$ applies to all the statistics within each
\alignmentSS{}{}.
\change{
However, the expectations for statistics that estimate gross diversity (e.g.,
$\pi$ and $\theta_W$) are not independent of sample size for structured
populations (e.g., the divergent pairs of populations modeled by \msb), and
other statistics are not independent of sample size in general (e.g.,
$SD(\pi-\theta_W)$).
Second, and more importantly, having the same expectation does not ensure
random variables are exchangeable.
Rather, for variables to be exchangeable their marginal distributions must be
the same (i.e., they must be identically distributed).
None of the summary statistics used by \msb, including $\pi_b$, are identically
distributed when there is any variation among taxa or loci in the
(1) numbers of sequences (\popSampleSize{1}{} and \popSampleSize{2}{}),
(2) length of alignments,
(3) numbers of loci (\nloci{}),
(4) \hky model parameters (\hkyModel{}{}),
(5) mutation-rate multipliers (\mutationRateScalarConstant{}{}),
or
(6) ploidy multipliers (\ploidyScalar{}{}).
Whenever such variation is present (i.e., nearly all empirical applications),
the taxon-specific summary statistics \alignmentSS{}{} are not exchangeable,
and the reshuffling of the summary statistic vectors is not mathematically
valid.

The magnitude of the affect of this violation of exchangeability is not known.
Huang et al.\ \cite{Huang2011} demonstrated that the reordering of the summary
statistic vectors can greatly increase the method's tendency to infer a single
divergence event.
By definition, if the summary statistic vectors were exchangeable, the
reordering would not change the likelihood or posterior (barring sampling
error).
Thus, the results of Huang et al.\ \cite{Huang2011} suggest the reordering of
the statistics is potentially introducing sizeable error to the analysis.
% Thus, the theoretical basis for this reshuffling is questionable.
% It may be justifiable as an additional coarsening of the data,
% by removing the information associated with the identity of the
% population pairs.
% However, I do not provide a formal proof that this re-ordering of the pairs
% does not introduce bias.
% One can imagine situations in which the sampling intensity (i.e., the
% number of gene copies, loci, and locus length) is highly skewed across
% pairs and/or there is large heterogeneity in mutational processes
% (e.g., \hkyModel{}{}, \mutationRateScalarConstant{}{}) among the pairs.
% In such cases, it seems possible that discarding this information could
% cause bias.
% Furthermore, given that (1) part of the motivation for re-ordering by $\pi_b$
% is to minimize Euclidean distances between simulated datasets in which the true
% model has a single divergence event \cite{Huang2011}, and (2) the reordering
% increases the model's tendency to infer a single divergence event
% \cite{Huang2011}, it seems possible that this approach could be biasing the
% method \cite{Oaks2012}.
}

@@

For each set of 50,000 simulated datasets, I used the posterior estimates
to assess the model-choice behavior of each model.
I did this by assigning the 50,000 estimates of the posterior probability
of one-divergence event to 20 bins of width 0.05, and plotted
the estimated posterior probability of each bin against the proportion of
replicates in that bin with a true value consistent with one divergence
event \cite{Huelsenbeck2004,Oaks2012}.
\change{
Ideally, the estimated posterior probability of the one-divergence model
should estimate the probability that the one-divergence model is correct.
For large numbers of simulation replicates, the proportion of the replicates in
each bin for which the one-divergence model is true will approximate the
probability that the one-divergence model is the correct model.
Thus, if the method has the desirable behavior such that the estimated
posterior probability of the one-divergence model is an unbiased estimate of
the probability that the one-divergence model is correct, the points should
fall near the identity line.
For example, let us say the method estimates a posterior probability of 0.90
for 1000 datasets simulated from the prior.
If the method is accurately estimating the probability that the one-divergence
model is correct given the data, then the one-divergence model should be the
true model in approximately 900 of the 1000 replicates.
Any trend away from the identity line indicates the method is biased in the
sense that it is not accurately estimating the probability that the
one-divergence model is the correct model.

@@

\subsection{Validation analyses: Ordered divergence models}
The results show that the method performs similarly when sampling over ordered
models of divergence
(Figures S\ref{figValidationAccuracyDPPOrderedDPPOrdered} \&
S\ref{figValidationModelChoiceDPPOrderedDPPOrdered}).
This suggests that the method is not adversely affected by the increase
in the number of possible discrete models (from 22 unordered to 4140 ordered
models) when there are eight pairs of populations.
\change{
This is encouraging, because, as discussed above, estimating unordered models
of divergence by shuffling the summary statistic vectors calculated from the
sequence alignments is not valid for most empirical datasets.
Given these results, estimation of unordered divergence models should be
avoided for empirical applications of the method.}

@@

\section{Discussion}
My results demonstrate that using alternative priors on parameters and
divergence models improved the behavior of the \msb model.
In the new implementation, model-choice estimation is more accurate and shows
greater robustness to model violations
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
The original model is very sensitive to violations and, when present,
strongly over-estimates the probability of one-divergence event shared
across all taxa
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).
\change{
When more appropriate priors are used for divergence-time and demographic
parameters, and either a Dirichlet-process or uniform prior applied across
divergence models, the model is less sensitive to violations, and, when
violations do cause bias, the method is conservative and tends to underestimate
the probability of models with temporally clustered divergences
(Figure~\ref{figValidationModelChoicePsi} \&
\ref{figValidationModelChoiceOmega}).}

@@

\change{
I also urge caution when using \dppmsbayes due to the lack of theoretical
validation of Bayesian model choice when the full data are replaced by summary
statistics that are insufficient for discriminating across models under
comparison \cite{Robert2011}, which is certainly the case here.
Robert et al.\ \cite{Robert2011} demonstrated that ABC estimates of model
posterior probabilities can be inaccurate when such across-model insufficient
statistics are used.

Given all of these caveats, I encourage investigators to view this method as a
means of exploring their data for general temporal patterns of divergences
across taxa, rather than a rigorous means of evaluating hypotheses.
As recommended by Oaks et al.\ \cite{Oaks2012}, any results from the method
should be accompanied by
(1) analyses under a variety of priors to assess the assumptions underlying
model inference and the prior sensitivity of the results, and
(2) simulation-based power analyses to provide insight into the temporal
resolution of the method.
Both approaches are important to help guide the interpretation of results.
}

@@

\section{Conclusions}
\change{
I introduced a new model for estimating shared divergence histories across taxa
from DNA sequence data within an approximate-Bayesian model-choice framework.
}
The new method, \dppmsbayes, takes a non-parametric approach to the problem by
using a Dirichlet-process prior on the temporal distribution of divergences
across taxa.
The new method shows improved robustness, accuracy, and power compared to the
existing method, \msb.
\change{
Compared to \msb, the new approach better estimates posterior uncertainty,
which greatly reduces the chances of incorrectly estimating biogeographical
scenarios of shared divergence events.
This is important, because models of shared divergence events are often of
particular interest to researchers who employ these methods.
}
This new tool will allow evolutionary biologists to better leverage comparative
genetic data to assess the affects of regional and global biogeographical
processes on biodiversity.
